---
title: "Stopover Annotation Tutorial"
author: "Julie Mallon"
date: "5/22/2020"
output: html_document
---

Using migration annotated data, annotate stopovers using FPT or non-sedentary stopovers with a nearest neighbor approach. Then run code that removes start or end points that are clearly movements.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

wd <- "C:/Users/Julie/Box/R/Chapter 1"
prjdta <- file.path(wd,"Data")
prjscript <- file.path(wd,"Scripts")
prjfuns <- file.path(wd,"Functions")
prjresults <- file.path(wd,"Results")
```

```{r}
source(file.path(prjfuns, "chapter 1 functions.r"))
#source(file.path(prjfuns, "migration annotation funs.r"))
source(file.path(prjfuns, "stopover annotation funs.r"))

require(adehabitatLT)
require(ggplot2)
require(dplyr)
library(lubridate)
library(data.table)
library(geosphere)

```

Prep Data to annotate stopovers
```{r}
birds_annotated <- read.csv(file.path(prjdta,"tuvu_mig_annotated.csv"), stringsAsFactors = F)
birds_annotated[which(birds_annotated$individual.local.identifier ==  "Artful_Dodger" & 
                        birds_annotated$date>"2018-09-12" & 
                        birds_annotated$date<"2018-09-15"),]$mig.state <- "Non-mig"
birds_annotated$ID_mig <- set.ID.pop(birds_annotated,
                                     birds_annotated$mig.state)
migration_data <- birds_annotated[which(birds_annotated$mig.state=="Mig"),]
example_data <- migration_data[which(migration_data$ID_mig==4),]
save(example_data,file=paste0(prjdta,"tuvu_migration_data.robj"))

migration_data$timestamp <- as.POSIXct(
  strptime(
    migration_data$timestamp,
    "%Y-%m-%d %H:%M:%S"),
  usetz=F)
migration_data$study.local.timestamp <- as.POSIXct( 
  strptime(
    migration_data$study.local.timestamp,  
    "%Y-%m-%d %H:%M:%S"),
  usetz=F)

#remove sparse migrations
bird <- split(migration_data,migration_data$individual.local.identifier)
migdata2<- lapply(bird, function(x){
  x<-x[order(x$study.local.timestamp),]
  x<-data.table(x)
  DT<-x[,.(duration=as.numeric(diff(range(study.local.timestamp))),
          n=.N),by=ID_mig]
  DT<-DT[,.(n/duration),by=ID_mig]  
  migs.keep<-DT[V1>10,ID_mig]
  x<-subset(x,ID_mig %in% migs.keep)
})

migdata3 <- rbindlist(migdata2)
#mig_ids_rm <- c(481, 436, 428)
#migdata3 <- migdata3[which(!(migdata3$ID_mig %in% mig_ids_rm))] 
```


```{r}



migs <- split(migdata3,
              migdata3$ID_mig)

# fix one migration annotation


mig_stops1 <- lapply(migs[1:20], function(x){
  k = x$ID_mig[1]

  #prevent issues due to gaps in track by interpolating missing points
  mig_nogap <- interpolate_gappy_track(x$X,
                                       x$Y,
                                       as.POSIXct(
                                         x$study.local.timestamp))
  
  df.traj <- as.ltraj(data.frame(mig_nogap$x,
                               mig_nogap$y),
                    as.POSIXct(mig_nogap$time), 
                    id=k)
  
  #calculate fpt
  df.fpt <- fpt_nogap(df.traj)
  myfpt <- data.frame(time = mig_nogap$time, 
                      fpt = df.fpt[[1]]$r1/60/60)
  
  #find threshold
  threshold <- find_threshold(myfpt, mig_nogap)
  
  #pull out likely stopovers
  stops <- subset(myfpt, fpt>threshold+1)
  
  mig_roost <- roost_times(x, mig_nogap)
  mig_nogap$hr <- lubridate::hour(mig_nogap$time)
  mig_nogap$roost <- ifelse(mig_nogap$hr>mig_roost[2] & mig_nogap$hr < mig_roost[1], 0, 1)
    
  
  if(nrow(stops)>0){
    mig_nogap$stopover <- ifelse(mig_nogap$time %in% stops$time,
                               1,0)
    mig_nogap$stop.id<- set.ID.ind(mig_nogap, mig_nogap$stopover)

    x2 <- mig_nogap %>% 
      group_by(stop.id,stopover,roost) %>%
      summarize(dur=length(OBJECTID))
    
    x2 <- subset(x2,stopover==1)
    x2 <- x2  %>% 
      group_by(stop.id) %>%
      mutate(p=dur/sum(dur))
    
    y<-as.data.frame(subset(x2, roost==0 & p>.2 & dur>2))
    
    mig_nogap$stopover <- ifelse(mig_nogap$stop.id %in%
                                   as.numeric(y[,1]),
                                 1,0)
    with(mig_nogap, 
         ScanTrack4(time, x,y,
                    col=as.factor(stopover),
                    main=paste(k,"no gap"))) 
    plot(myfpt,type="l")
    abline(h=threshold)
    
  #mig2<-mig2[,c(1:22,26:28)]

  }else{
    mig_nogap$stopover <- 0
    mig_nogap$stop.id<- set.ID.ind(mig_nogap, mig_nogap$stopover)
    #mig2 <- mig[,c(21,1:20,22:25)]
  }
  
  mig2 <- merge(x, 
              mig_nogap,
              by.x="study.local.timestamp",
              by.y="time",
              all.x=T,
              all.y=F)
  
  #check that no stopovers were missed with a buffer check
  mig.sp <- mig2
  coordinates(mig.sp) <- c("location.long","location.lat")
  dm<-data.frame(coordinates(mig.sp),
                 time=mig.sp$study.local.timestamp, 
                 X=rowSums(distm (coordinates(mig.sp)[,1:2], 
                                fun = distHaversine) / 1000 <= 15)) # number of points within distance 10 km
  mig2[which(dm$X>30),'stopover']<-1
  mig2<- mig2[-nrow(mig2),]
  mig2$stop.id<- set.ID.ind(mig2, mig2$stopover)
  with(mig2, ScanTrack(study.local.timestamp,
                       X,
                       Y,
                       col=as.factor(stopover),
                       main=paste(k,"15 km buffer")))

  return(mig2)
  }
)

  
# left off here

 stops_summary_table(mig_stops1[[3]]) 

 
 #save(stop_fpt2,file=paste0(prjdta,"stopover_fpt2.robj"))
#save(adedf,file=paste0(prjdta,"stopover_adehab_data.robj"))
#save(vars.table,file=paste0(prjdta,"stopover_annot_variables.robj"))
mig_stops1 <-   rbindlist(mig_stops1)
mig_stops1$study.local.timestamp <- as.character(mig_stops1$study.local.timestamp)
fwrite(mig_stops1, file.path(prjdta, "stop_ann1.csv"))

```

#left off having just run this for all birds. Need to check robustness against previous method and then categorize the stops

```{r}


stops_mia <- c(477,453, 445, 385, 363 )

nm_overlap_stop <- c(469, 361 )

too_short <- c(462, )

too_many <- c(457, 378, 368, 366, 359, 357, 354)
```

```{r, eval = F}
# x<- mig_stops1[[2]] 
# x$stop.start <- c(ifelse(diff(x$stop.id)== 1, 1, 0), 0) 
# y <- c(1,which(x$stop.start==1))
# x[y]
# x %>% group_by(stop.id) %>% 
#   summarize()

fix_mig.state <- function(x){
  x$days.cont <- as.numeric(x$date - x$date[1])
  x$ID_mig <- set.ID(x)
  x2 <- x %>%
    group_by(ID_mig, mig.state) %>%
    summarize(dday = max(days.cont) - min(days.cont)) %>%
    ungroup()
  
  x.table.sub <- which(x2$mig.state=="Mig")
  min.dday <- which.min(x2[x.table.sub,]$dday)
  max.ids <- round((length(which(table(x$year,
                                       month(x$study.local.timestamp))
                                 >0))+4)/3)
  repeat{
    if(x2[x.table.sub,]$dday[min.dday]<15){
      fix.ms <- x2$mig.state[x.table.sub[min.dday]]
      fix.ms <- ifelse(fix.ms == "Non-mig", "Mig", "Non-mig")
      fix.id <- x2$ID_mig[x.table.sub[min.dday]]
      x$mig.state[which(x$ID_mig == fix.id)] <- fix.ms
      x$ID_mig <- set.ID(x)
    }
    x2 <- x %>%
      group_by(ID_mig, mig.state) %>%
      summarize(dday = max(days.cont) - min(days.cont)) %>%
      ungroup()
    x.table.sub <- which(x2$mig.state=="Mig")
    min.dday <- which.min(x2[x.table.sub,]$dday)
    if(max(x$ID_mig) <= max.ids & min(x2$dday>4)){
      break
    }
  }
  #ScanTrack(x$study.local.timestamp, 
  #x$X, x$Y, col = as.factor(x$ID_mig))
  #print(x$individual.local.identifier[1])
  return(x)
}   
```

```{r, eval =F}
load(paste0(prjdta,"stopover_fpt2.robj"))
load(paste0(prjdta,"stopover_adehab_data.robj"))
stop_fpt2<-as.data.frame(stop_fpt2)
stop_fpt2<-stop_fpt2[order(stop_fpt2$ID_migration,stop_fpt2$local.time),]
stop_fpt2$stop.id<- cumsum(ifelse(c(1,diff(stop_fpt2$stopover)),1,0))
stop_fpt2$tag.local.identifier<-as.factor(stop_fpt2$tag.local.identifier)
DF<-merge(stop_fpt2,adedf,by.x=c("local.time","X","Y","ID_migration"),by.y=c("date","x","y","id"),all=FALSE)

save(DF,file=paste0(prjdta,"stopover_annotations.robj"))
```

```{r check stopovers match earlier annotations, eval = F}
load(file.path(wd,"2017/Data", "stopovers_with_all_weather.robj"))
DF_old <- stops_full_weather 

#df_old_inds <- split(DF_old, DF_old$individual.local.identifier)
DF_old %>% 
  group_by(individual.local.identifier, halfyr, stopover, day, stop.id) %>%
  summarize(n.days = n()) %>%
  group_by(individual.local.identifier, halfyr, stopover) %>%
  summarize(n.stop = length(unique(stop.id)), n.days = n() ) %>%
  filter(stopover==1) %>% View()

new_out <- rbindlist(mig_stops1) %>% 
  group_by(individual.local.identifier, halfyr, stopover, day, stop.id) %>%
  summarize(n.days = n()) %>%
  group_by(individual.local.identifier, halfyr, stopover) %>%
  summarize(n.stop = length(unique(stop.id)), n.days = n() ) %>% 
  filter(stopover==1)


DF_old[which(DF_old$individual.local.identifier=="Desert_Rat"),] -> temp
temp[which(temp$halfyr=="2015 B"),] -> temp
with(temp,
ScanTrack(study.local.timestamp,
X,
Y,
col = as.factor(stop.id)))

with(mig_stops1[[62]],
ScanTrack(study.local.timestamp,
X,
Y,
col = as.factor(stop.id)))
```

May need to return to the stopover annotation - go with results for now (april 14)


Next,'stopover analyses june 2017.rmd'

do analyses with and without definite feeding stops (high movement and long duration)
