---
title: "Stopover Annotation Tutorial"
author: "Julie Mallon"
date: "5/22/2020"
output: html_document
---

Using migration annotated data, annotate stopovers using FPT or non-sedentary stopovers with a nearest neighbor approach. Then run code that removes start or end points that are clearly movements.
    
*Note:* These functions assume names are the same as Movebank data frames.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

wd <- getwd()

source(file.path(wd, "stopover annotation tutorial functions.r"))

library(adehabitatLT)
library(dplyr)
library(lubridate)
library(geosphere)     # distm function
library(suncalc) 
```

```{r required functions and packages, eval = F}
source(file.path(wd, "stopover annotation tutorial functions.r"))

library(adehabitatLT)
library(dplyr)
library(lubridate)
library(geosphere)     # distm function
library(suncalc)       # required for roost times function
```

### Load and prep data to annotate stopovers
```{r}
load(file=file.path(wd,"example_data.robj"))

# fix timestamp formats
example_data$timestamp <- as.POSIXct(
  strptime(
    example_data$timestamp,
    "%Y-%m-%d %H:%M:%S"),
  usetz=F)
example_data$study.local.timestamp <- as.POSIXct( 
  strptime(
    example_data$study.local.timestamp,  
    "%Y-%m-%d %H:%M:%S"),
  usetz=F)

```

### Stopover annotation function
```{r}

annotate_stops_fpt <- function(x){
  # run for a single individual, single migration
  # to run for multiple individuals or migrations, use lapply
   
  # prevent issues due to gaps in track by interpolating missing points
  mig_nogap <- interpolate_gappy_track(x$X,           # must be in UTMs and not degrees!
                                       x$Y,
                                       as.POSIXct(
                                         x$study.local.timestamp))
  
  df.traj <- as.ltraj(data.frame(mig_nogap$x,         # format data for adehabitat
                               mig_nogap$y),
                    as.POSIXct(mig_nogap$time), 
                    id=x$individual.local.identifier[1])
  
  
  df.fpt <- fpt(df.traj, r = 20000)                   # calculate FPT, check that r value works for your dataset
  myfpt <- data.frame(time = mig_nogap$time,          # append FPT to a new data frame with time
                      fpt = df.fpt[[1]]$r1/60/60)
  
  
  threshold <- find_threshold(myfpt, mig_nogap)       # find threshold to identify stopovers
  
  
  stops <- subset(myfpt, fpt > threshold + 1)         # pull out likely stopovers
  
  mig_roost <- roost_times(x, mig_nogap)
  mig_nogap$hr <- lubridate::hour(mig_nogap$time)
  mig_nogap$roost <- ifelse(mig_nogap$hr>mig_roost[2] & 
                              mig_nogap$hr < mig_roost[1], 
                            0, 1)
    
  
  if(nrow(stops)>0){                                 # if some stops have been identified
    mig_nogap$stopover <- ifelse(mig_nogap$time %in% stops$time,
                               1,0)
    mig_nogap$stop.id<- set.ID.ind(mig_nogap, mig_nogap$stopover)

    x2 <- mig_nogap %>% 
      group_by(stop.id,stopover,roost) %>%
      summarize(dur=length(OBJECTID))
    
    x2 <- subset(x2,stopover==1)
    x2 <- x2  %>% 
      group_by(stop.id) %>%
      mutate(p=dur/sum(dur))                         # proportion of stopover that is spent roosting
                                                     # Up to the user to decide whether this helps with selection
    
    y <- as.data.frame(
      subset(x2,
             roost==0 & p>.2 & dur>2))               # I find 0.2 works for this data (typical roost 6pm - 8am) 
    
    mig_nogap$stopover <- ifelse(mig_nogap$stop.id %in%
                                   as.numeric(y[,1]),
                                 1,0)
    with(mig_nogap, 
         ScanTrack4(time, x,y,
                    col=as.factor(stopover),
                    main="no gap"))                   # Scantrack plot to see where stopovers have been annotated
    plot(myfpt, type="l")                             # FPT plot
    abline(h=threshold)
    

  }else{
    mig_nogap$stopover <- 0
    mig_nogap$stop.id<- set.ID.ind(mig_nogap, mig_nogap$stopover)
  }
  
  mig2 <- merge(x, 
              mig_nogap,
              by.x="study.local.timestamp",
              by.y="time",
              all.x=T,
              all.y=F)
  
  #check that no stopovers were missed with a buffer check
  mig.sp <- mig2
  coordinates(mig.sp) <- c("location.long", "location.lat")
  dm<-data.frame(coordinates(mig.sp),
                 time=mig.sp$study.local.timestamp, 
                 X=rowSums(distm (coordinates(mig.sp)[,1:2], 
                                fun = distHaversine) / 1000 <= 15)) # number of points within distance 15 km
  mig2[which(dm$X > 30),'stopover']<-1
  mig2<- mig2[-nrow(mig2),]
  mig2$stop.id<- set.ID.ind(mig2, mig2$stopover)
  with(mig2, ScanTrack(study.local.timestamp,
                       X,
                       Y,
                       col=as.factor(stopover),
                       main="15 km buffer"))

  return(mig2)
  }
```


### Run annotation function
```{r}
example_annotated <- annotate_stops_fpt(example_data)      # the above function
```

### Optional summary table of stopovers
```{r}
stops_summary_table(example_annotated)                     # optional list and plots summarizing stopovers

```

End code.

